\textbf{How to measure maximally precise radial velocities from a set of spectra:\\ The time-independent spectrum case}

### 2019-08-26 paper/eprv.tex
cross correlation

The most optimistic assumptions we can possibly make are the following.
\begin{itemize}
\item The spectrograph is perfectly calibrated in a wavelength sense.
\item The spectrograph is consistently and repeatably calibrated in a
  flux sense. That is, the spectrum does not need to be perfectly
  continuum-normalized or perfectly flux-normalized, but it is
  calibrated near-perfectly \emph{consistently} across epochs (exposures).
  We will break this assumption in our experiments, below.
\item There are no residual telluric absorptions nor any residual sky
  emission in the spectra. That is, the sky is near-perfectly calibrated
  and removed.
  We will also break this assumption in our experiments, below.
\item The \emph{true} spectrum (that is, the latent spectrum that
  would be observed if the data were far, far better) of the star is
  known at better accuracy and precision than the main noise sources.
  We will break this assumption too.
  The whole point of this paper is to ask whether we can create a
  good-enough proxy for the true spectrum.
\item There is no time dependence (no epoch-to-epoch variations) in the
  stellar spectrum.
\item All noise sources contributing to the spectral data---which
  include at a minimum photon noise, read noise, residual telluric
  issues, residual wavelength-calibration issues, and residual stellar
  spectral variability---are, when summed together, indistinguishable
  from being zero-mean, Gaussian, and additive.
\item The complete variance tensor of the total noise on the spectrum is known, and
  known correctly. That is, the variance of the noise at every pixel
  is known, as is the covariance between nearby pixels, and in the
  same \emph{true} sense that the true spectrum of the star is known.
\end{itemize}
It is worthy of note that every formula (that we know) in the
literature for maximum \RV\ precision (\todo{cite Butler, Bouchy,
  Figueira, Lovis}) makes all of these same assumptions,
plus additional \emph{even more restrictive assumptions}.
Under these disturbingly unrealistic assumptions, the \CRLB\ becomes
\begin{equation}\label{eq:crlb}
\frac{1}{\sigma_v^2} \leq \left[\frac{\dd f}{\dd v}\right]\T\cdot C^{-1}\cdot\left[\frac{\dd f}{\dd v}\right]
\end{equation}
where the derivatives are of the true spectrum $f$ with respect to
velocity $v$ (which can, in turn, be thought of as the derivative of a Doppler
operator with respect to velocity acting on the true spectrum; CITE WOBBLE), the
derivatives are column vectors, and $C^{-1}$ is the inverse of the covariance matrix
describing the true variance and covariances of the noise process.

@@

@@

HOGG: OPEN by giving the LF and what it's optimum is, in the simplest case.
Parameterize $f(v)$ where $v$ is the \RV.

Most methods for radial-velocity estimation involve some kind of
cross-correlation of the data with a template.
This might be surprising, since (as we noted above) bound-saturating
estimators will be maximum-likelihood estimators.
However, under some simple assumptions, cross-correlation optima are
also maximum-likelihood optima.

For example, when the likelihood is Gaussian, the log-likelihood
contains a chi-squared-like term
\begin{equation}
-\frac{1}{2}\,[y - f]\T\cdot C^{-1}\cdot [y - f] \quad ,
\end{equation}
where $y$ is a column vector of the data, $f$ is the model spectrum,
and $C$ is the covariance matrix describing the noise variance.
In the case of independent pixel noise this just becomes an
inverse-variance weighted sum of squared residuals (chi-squared).
If we expand this quadratic form, it becomes
\begin{equation}
{}-\frac{1}{2}\,y\T\cdot C^{-1}\cdot y + y^T\cdot C^{-1}\cdot f 
-\frac{1}{2}\,f\T\cdot C^{-1}\cdot f \quad.
\end{equation}
That is, the chi-squared term factors into a term that is the square
of the model, a term that is the square of the data, and a term that
is a cross-correlation.
As long as the amplitude of $f$ doesn't depend on velocity, and the
$C$ tensor is ``smooth'' in certain respects, the dependence of this
chi-squared term on velocity will be dominated by the
cross-correlation.
For this reason, we will use a variant of cross-correlation
optimization as our radial-velocity estimator.

>>>>

The most optimistic assumptions we can possibly make are the following.
\begin{itemize}
\item The spectrograph is perfectly calibrated in a wavelength sense.
\item The spectrograph is consistently and repeatably calibrated in a
  flux sense. That is, the spectrum does not need to be perfectly
  continuum-normalized or perfectly flux-normalized, but it is
  calibrated near-perfectly \emph{consistently} across epochs (exposures).
  We will break this assumption in our experiments, below.
\item There are no residual telluric absorptions nor any residual sky
  emission in the spectra. That is, the sky is near-perfectly calibrated
  and removed.
  We will also break this assumption in our experiments, below.
\item The \emph{true} spectrum (that is, the latent spectrum that
  would be observed if the data were far, far better) of the star is
  known at better accuracy and precision than the main noise sources.
  We will break this assumption too.
  The whole point of this paper is to ask whether we can create a
  good-enough proxy for the true spectrum.
\item There is no time dependence (no epoch-to-epoch variations) in the
  stellar spectrum.
\item All noise sources contributing to the spectral data---which
  include at a minimum photon noise, read noise, residual telluric
  issues, residual wavelength-calibration issues, and residual stellar
  spectral variability---are, when summed together, indistinguishable
  from being zero-mean, Gaussian, and additive.
\item The complete variance tensor of the total noise on the spectrum is known, and
  known correctly. That is, the variance of the noise at every pixel
  is known, as is the covariance between nearby pixels, and in the
  same \emph{true} sense that the true spectrum of the star is known.
\end{itemize}
It is worthy of note that every formula (that we know) in the
literature for maximum \RV\ precision (\todo{cite Butler, Bouchy,
  Figueira, Lovis}) makes all of these same assumptions,
plus additional \emph{even more restrictive assumptions}.
Under these disturbingly unrealistic assumptions, the \CRLB\ becomes
\begin{equation}\label{eq:crlb}
\frac{1}{\sigma_v^2} \leq \left[\frac{\dd f(v)}{\dd v}\right]\T\cdot C^{-1}\cdot\left[\frac{\dd f(v)}{\dd v}\right]
\end{equation}
where the derivatives are of the true spectrum $f(v)$ with respect to
radial velocity $v$ (which can, in turn, be thought of as the derivative of a Doppler
operator with respect to velocity acting on the true spectrum; CITE WOBBLE), the
derivatives are column vectors, and $C^{-1}$ is the inverse of the covariance matrix
describing the true variance and covariances of the noise process.

@@

\CRLB-saturating estimators are (or are equivalent to)
maximum-likelihood estimators.
When the likelihood is Gaussian, the log-likelihood
is a sum of a quadratic chi-squared-like term and a determinant term
\begin{equation}
\ln L = -\frac{1}{2}\,[y - f(v)]\T\cdot C^{-1}\cdot [y - f(v)] - \frac{1}{2}\,||2\pi\,C|| \quad ,
\label{eq:lf}
\end{equation}
where $y$ is a column vector of the data, $f(v)$ is the model spectrum,
which depends on the \RV\ $v$,
and $C$ is the covariance matrix describing the noise variance.
In the case of independent pixel noise the chi-squared term just becomes an
inverse-variance weighted sum of squared residuals (chi-squared).
The maximum-likelihood estimator for $v$ is the value of $v$ which maximizes
the log-likelihood.
Since only the chi-squared term depends on $v$, maximizing the likelihood
means minimizing chi-squared.

Most methods for radial-velocity estimation aren't explicitly maximizing
any likelihood.
Instead, they involve some kind of
cross-correlation of the data with a template.
This might be surprising, since (as we noted above) bound-saturating
estimators will be maximum-likelihood estimators.
However, under simple assumptions, cross-correlation optima are
also maximum-likelihood optima. The argument is as follows:
If we expand the quadratic form (chi-squared term) in the
likelihood in equation~(\ref{eq:lf}), it becomes
\begin{equation}
{}-\frac{1}{2}\,y\T\cdot C^{-1}\cdot y + y^T\cdot C^{-1}\cdot f(v) 
-\frac{1}{2}\,f(v)\T\cdot C^{-1}\cdot f(v) \quad.
\end{equation}
That is, the chi-squared term factors into a term that is the square
of the model, a term that is the square of the data, and a term that
is a cross-correlation.
As long as the amplitude of $f$ doesn't depend on velocity, and the
$C$ tensor is ``smooth'' in certain respects, the dependence of this
chi-squared term on velocity will be dominated by the
cross-correlation.
For this reason, we can use a variant of cross-correlation
optimization as a radial-velocity estimator.

@@
